---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nestr

<!-- badges: start -->
<!-- badges: end -->


```{r}
library(nestr)
```

The main purpose of `nestr` R-package is to build nested (or hierarchical) structures with output as a list or a data frame. The syntax is particularly useful when the child units are unbalanced across the parental levels. 


```{r}
nest_in(c("a", "c", "a", "b"), # parental vector
        2 ~ 3, # level 2 has 3 children
        . ~ 1) # the remaining levels have 1 child 
```
The parental vector may be a factor with different ordering of the levels.

```{r}
nest_in(factor(c("a", "c", "a", "b"), levels = c("a", "c", "b")), # parental vector
        2 ~ 3, # level 2 has 3 children
        . ~ 1) # the remaining levels have 1 child 
```

Or you can refer the parental level by character.

```{r}
nest_in(c("a", "c", "a", "b"), # parental vector
        "b" ~ 3, # "b" has 3 children
        . ~ 1) # the remaining levels have 1 child 
```

A more interesting example.

```{r}
nest_in(c("Math", "Science", "Economics", "Art"), 
               "Science" ~ 2,
    c("Art", "Math") ~ 10,
                    . ~ 3,
        prefix = "student-",
        leading0 = 4)
```

## `edibble::nested_in` and `nestr::nest_in`

The syntax of `nestr` is similar to `edibble::nested_in` which is used within the [`edibble::set_units`](https://edibble.emitanaka.org/reference/set_units.html) to construct nested (or hierarchical) structures. 


Unlike `edibble::nested_in`, the `nestr::nest_in` returns a data frame. You may also notice that the name of the function is different although both share virtually the same syntax. One way to remember the differences in function name is that `nested_in` is for `edibble` and is meant for construction of the experimental design (notice the "ed" in the function name). When the verb is written in present tense (i.e. `nest_in`), it's part of `nestr` and your focus is to create a structure in the present.


## `amplify`

The `dplyr::mutate` function modifies, creates or deletes columns but doesn't alter the number of rows. The `nestr::amplify` function can create new columns which generally increase (or amplify) the size of the row dimension. The columns that were amplified as a result of the created column will be duplicated. If you are familiar with gene replication process then you can recall these functions in those terms. An amplified gene is just a duplication of the original. A mutated gene modifies the original state.

```{r}
df <- data.frame(country = c("AU", "NZ", "JPN", "CHN", "USA")) %>% 
  amplify(soil = nest_in(country, 
                            "AU" ~ 10,
                            "NZ" ~ 8,
                               . ~ 5,
                           prefix = "sample",
                           leading0 = TRUE),
          rep = nest_in(soil, 
                          1:3 ~ 3, # first 3 samples have 3 technical rep
                            . ~ 2)) # remaining have two rep

tibble::as_tibble(df)
```

## `tidyverse`

For those who want to stay within the tidyverse framework, this is in fact doable just using `dplyr` and `tidyr`. 

The semantics are less direct, however and chaining of nesting structure is cumbersome. For example, see the equivalent example from before below.


```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
data.frame(country = c("AU", "NZ", "JPN", "CHN", "USA")) %>% 
  mutate(soil = case_when(country=="AU" ~ list(sprintf("sample%.2d", 1:10)),
                          country=="NZ" ~ list(sprintf("sample%.2d", 1:8)),
                               TRUE ~ list(sprintf("sample%.2d", 1:5)))) %>% 
  unnest_longer(soil) %>% 
  mutate(rep = case_when(soil %in% c("sample01", "sample02", "sample03") ~ list(1:3),
                           TRUE ~ list(1:2))) %>% 
  unnest_longer(rep)
```

The intent I think is more clear from the above `amplify` example. It's a personal preference so use what suits your own situation!

