---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nestr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->


```{r}
library(nestr)
```

## `edibble::nested_in` and `nestr::nest_in`

The syntax of `nestr` is the similar to `edibble::nested_in` which is used within the [`edibble::set_units`](https://edibble.emitanaka.org/reference/set_units.html) to construct nested (or hierarchical) structures. 

The syntax proves useful in situations where the child units are unbalanced across the parental levels. 

```{r}
nest_in(1:3, # parental vector
        1 ~ 3, # level 1 has 3 children
        . ~ 1) # the remaining levels have 1 child 
```




Unlike `edibble::nested_in`, the `nestr::nest_in` returns a data frame. You may also notice that the name of the function is different although both share similar syntax. One way to remember the differences in function name is that `nest`**`ed`**`_in` is for `edibble` and is meant for construction of the experimental design. When the verb is written in present tense (i.e. `nest_in`), it's part of `nestr` and your focus is to create a structure in the present.



A more interesing example.

```{r}
farmer <- c("John", "Jane", "Ann", "Thomas", "Helen")
nest_in(farmer, 
               "John" ~ 2,
    c("Ann", "Helen") ~ 10,
                    . ~ 3,
        name = "chick",
        prefix = "chick-",
        leading0 = TRUE,
        min_ndigits = 4)
```


## `amplify`

Not implemented yet.

The `dplyr::mutate` function modifies, creates or deletes columns but doesn't alter the number of rows. The `nestr::amplify` function can create new columns which generally increase (or amplify) the size of the row dimension. The columns that were amplified as a result of the created column will be duplicated. If you are familiar with gene replication process then you can recall these functions in those terms. An amplified gene is just a duplication of the original. A mutated gene modifies the original state.

```{r, eval = FALSE}
data.frame(country = c("AU", "NZ", "JPN", "CHN", "USA")) %>% 
  amplify(person = nest_in(country, 
                            "AU" ~ 20,
                            "NZ" ~ 10,
                               . ~ 100),
          child = nest_in(person, 
                          1:10 ~ 3,
                             . ~ 2))
```


## `tidyverse`

For those who want to stay within the tidyverse framework, this is in fact doable just using `dplyr` and `tidyr`. 

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
data.frame(parent = 1:3) %>% 
  mutate(child = case_when(parent==1 ~ list(1:3),
                       TRUE ~ list(1:2))) %>% 
  unnest_longer(child)
```

The semantics are less direct, however and chaining of nesting structure is cumbersome. For example, see the country example below.


```{r}
data.frame(country = c("AU", "NZ", "JPN", "CHN", "USA")) %>% 
  mutate(person = case_when(country=="AU" ~ list(1:20),
                            country=="NZ" ~ list(1:10),
                               TRUE ~ list(1:100))) %>% 
  unnest_longer(person) %>% 
  mutate(child = case_when(person %in% unique(person)[1:10] ~ list(1:3),
                           TRUE ~ list(1:2))) %>% 
  unnest_longer(child)
```

The intent I think is more clear from the above `amplify` example.

